---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
- By Yang Yuzhong

The activity.csv contains a total of 17,568 observation of the variables:

- **steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)

- **date**: The date on which the measurement was taken in YYYY-MM-DD format

- **interval**: Identifier for the 5-minute interval in which measurement was taken

## Loading required R packages

```{r loadpackages, echo=TRUE}
#library(dplyr)
```

## Loading and preprocessing the data

activity.csv is loaded using read.csv() function. No preprocessing is performed.

```{r loadingdata, echo=TRUE}
#activity<-read.csv("activity.csv")
```

The format of the loaded table is then changed to the following with select 
function from dply:

| date | interval | steps |
| ---- | -------- | ----- |
| 2012-10-01 | 0 | NA |
| ... | ... | ... |

```{r preprocessdata, echo=TRUE}
#activity<-activity %>% 
#select(date,interval,steps) %>%
#mutate(steps = ifelse(is.na(steps),0,steps))
```

## What is mean total number of steps taken per day?

### 1) Histogram of total number of steps taken per day

```{r histogram, echo=TRUE}
activity.totalsteps <- activity %>% 
    group_by(date) %>% 
    summarize(totalsteps = sum(steps, na.rm=TRUE))

#hist(activity.totalsteps$totalsteps,
#main="Distribution of total number of steps taken per day",
#xlab="Total number of steps taken per day (Grouped)",
#ylab="Frequency of groups")
```

### 2) Mean and median total number of steps taken per day

To substitute a paragraph with mean and median values (based on the coursera lesson)

```{r meanmedian, echo=TRUE}
mean(activity.totalsteps$totalsteps)

median(activity.totalsteps$totalsteps)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
